---
title: "HW 03 - Road traffic accidents"
author: "Jieming Xiao"
date: "07/14/2022"
output: github_document
---

## Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse)
library(dsbox)
```

## Exercises

### Exercise 1

There are `r nrow(accidents)` rows in the dataset. 

### Exercise 2

Each row represents a case of road traffic accident.

### Exercise 3

For weekends, most accidents happen during afternoon with the peak around 4pm. For weekdays, a lot of accidents happen during rushing hours around 8am and 4pm.


```{r accidents_week_of_day}
accidents_recode <-accidents %>%
  mutate(week = recode(day_of_week,
                       `Monday` = "Weekday",
                       `Tuesday` = "Weekday",
                       `Wednesday` = "Weekday",
                       `Thursday` = "Weekday",
                       `Friday` = "Weekday",
                       `Saturday` = "Weekend",
                       `Sunday` = "Weekend"))
ggplot(accidents_recode, aes(x = time, 
                  fill = severity)) +
  geom_density(adjust = 1, 
               alpha = 0.5
               ) +
    theme_minimal()+
 facet_wrap(~week,
            nrow =2) +
  labs(
    x = "Time of day",
    y = "Density",
    title = "Number of accidents throughout the day", 
    subtitle = "By day of week and severity",
    fill = "Severity",
  )

```

### Exercise 4

Most accidents happen on single carriageway when the road has frost or ice and when speed limit is low (below 30).The other road types and speed limits don't have a lot of accidents. 

```{r accidents_road}
ggplot(accidents_recode, aes(x = speed_limit, 
                  fill = road_surface)) +
  geom_density(adjust = 3, 
               alpha = 0.5
               ) +
    theme_minimal()+
 facet_wrap(~road_type,
            nrow =5) +
  labs(
    x = "Speed limit",
    y = "Density",
    title = "Number of accidents across speed limit", 
    subtitle = "By road surface",
    fill = "Road Surface",
  )
```
